---
title: "Nike Sales Data 2020-2021"
output: 
  html_document:
    toc: true
    toc_float: true
    number_section: true
    highlight: tango
    theme: readable
author: "Jamie Kerlin"
date: "2024-07-30"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```


# Overview

## Purpose

The purpose of this document is to clean and analyze the Nike Sales Dataset 2020-2021.

## Data Citation

This is a test dataset downloaded from Kaggle on 2024-07-30 https://www.kaggle.com/datasets/pavankota2020/nike-sales-dataset?resource=download

# Setup

## Load libraries

```{r}
package_names <- c("tidyverse", "here", "janitor", "vtable", "ggimage", "lubridate",
                   "car", "marginaleffects", "ggeffects", "olsrr", "ggcorrplot")

for(x in package_names){
  if(!x %in% rownames(installed.packages())) install.packages(x)
}

library(tidyverse)
library(here)
library(janitor)
library(lubridate)
library(vtable)
library(ggimage)
library(car)
library(marginaleffects)
library(ggeffects)
library(olsrr)
library(ggcorrplot)

```


## Load data

```{r}

data <- read_csv(here("Nike_Sales_Data",
                      "Data",
                      "Data_Raw",
                      "nike_data.csv"))

```

# Data exploration and cleaning

## Clean names

Change the column names to clean names using janitor package. Create unique row ID column

```{r}

data <- data %>%
  clean_names() %>%
  mutate(id = 1:n()) %>%
  select(id, everything())

```

## Explore data

First, check for basics such as missing values, the number of distinct values in each column, and the data class for each of the columns.
```{r}

# 9360 rows, 9 columns

# Check for any NAs
nas <- data %>%
  filter(!complete.cases(.))

# How many distinct values are in each of the columns?
data %>%
  map_dbl(n_distinct)

# What classes of data are in the dataset?
vtable(data)
  
# see if duplicated rows for all categories
dupes <- data %>%
  select(where(is.character)) %>%
  get_dupes()


# look at those duplicated rows- there are multiple rows for categories but no unique identifier for the products
# so the duplicates are OK
get_dupes <- left_join(dupes, data)
  
```


## Descriptives

Next, I will look at descriptives of those column values

What are the frequencies of the values in the columns?

```{r}

sumtable(data)

data_gg <- data %>%
  pivot_longer(cols = product:sales_method,
               names_to = "characteristic",
               values_to = "characteristic_value") %>%
  group_by(characteristic, characteristic_value) %>%
  summarise(ch_num = n())


ggplot(data_gg, mapping = aes(x = characteristic_value, y = ch_num)) +
  geom_col() +
  coord_flip() +
  facet_wrap(~characteristic, scales = "free")

data_gg_num <- data %>%
  pivot_longer(cols = price_per_unit:units_sold,
               names_to = "characteristic",
               values_to = "characteristic_value")


ggplot(data_gg_num, mapping = aes(x = characteristic_value)) +
  geom_histogram() +
  facet_wrap(~characteristic, scales = "free")


data %>%
  mutate(state = fct_infreq(state)) %>%
  ggplot(mapping = aes(x = state, fill = region)) +
    geom_bar() +
    coord_flip() +
    facet_wrap(~region, scales = "free")


```

Check invoice date

I want to make sure the date format is consistent in the invoice date column

```{r}

date_test <- data %>%
  mutate(year = str_sub(invoice_date, 7, 10),
         month = str_sub(invoice_date, 4, 5),
         day = str_sub(invoice_date, 1, 2))

date_test %>% count(year)
date_test %>% count(month)
date_test %>% count(day)

# ok, date formats are correct- use lubridate on the column to get final consistent format
data$invoice_date <- lubridate::dmy(data$invoice_date)
```

Box plots

```{r}
data_sales <- data %>%
  pivot_longer(cols = product:state,
               names_to = "cat_char",
               values_to = "cat_value") %>%
  pivot_longer(cols = price_per_unit:units_sold, 
               names_to = "num_char",
               values_to = "num_value") %>%
  group_by(cat_char, cat_value, num_char) %>%
  filter(cat_char != "state") %>%
  ungroup() %>%
  group_by(cat_char) %>%
  nest()
```

```{r}
output <- list()

for(y in 1:4){
plot <- data_sales %>%
  pluck("data", y) %>%
  ggplot(mapping = aes(x = cat_value, y = num_value)) + 
    geom_boxplot() + 
    coord_flip() +
    facet_wrap(~num_char, ncol = 3, scales = "free_x") 

output[[y]] <- plot
}

print(output)
```


Create a dataset to compare sales by categories

```{r}

data_sales <- data %>%
  pivot_longer(cols = product:state,
               names_to = "cat_char",
               values_to = "cat_value") %>%
  pivot_longer(cols = price_per_unit:units_sold, 
               names_to = "num_char",
               values_to = "num_value") %>%
  group_by(cat_char, cat_value, num_char) %>%
  summarise(mean_num_value = mean(num_value),
            total_num_value = sum(num_value)) %>%
  filter(cat_char != "state") %>%
  ungroup() %>%
  group_by(cat_char) %>%
  nest()


```


Compare average daily sales

```{r}
output <- list()

for(y in 1:4){
plot <- data_sales %>%
  pluck("data", y) %>%
  ggplot(mapping = aes(x = cat_value, y = mean_num_value)) + 
    geom_col() + 
    coord_flip() +
    facet_wrap(~num_char, ncol = 3, scales = "free_x") 

output[[y]] <- plot
}

print(output)
```

On average per day:  
  - Women and men's apparel is more expensive than footwear   
  - Women's apparel and men's street footwear sells the most units, women's athletic footwear sells the least  
  - Price per unit is lower in the south and midwest  
  - Most units sold in the west, southeast, and south  
  - Price per unit between retailers not very much  
  - Walmart sells the most units, Amazon sells the least  
  - Outlets have lower price per units  
  - Most sales from in-person  


Compare total sales over 2 years

```{r}
output <- list()

for(y in 1:4){
plot <- data_sales %>%
  pluck("data", y) %>%
  ggplot(mapping = aes(x = cat_value, y = total_num_value)) + 
    geom_col() + 
    coord_flip() +
    facet_wrap(~num_char, ncol = 3, scales = "free_x") 

output[[y]] <- plot
}

print(output)
```

Total throughout the two years:  
  - Women and men's apparel is more expensive than footwear     
  - Women's apparel and men's street footwear sells the most units, followed closely by men's athletic footwear  
  - Women's athletic footwear and men's apparel sell the fewest units  
  - Price per unit is way higher in west and northeast  
  - Most units sold in the west, northeast, and south    
  - Price per unit very different but because more units sold at some places  
  - Most units sold at West Gear, Foot Locker, and Sports Direct  
  - Most units sold online, fewest in-store  
  
  
What do the sales look like throughout the two years?

Make a dataset to look at sales by year, quarter, month
```{r}

data_time <- data %>%
  mutate(year = year(invoice_date),
         quarter = quarter(invoice_date),
         month = month(invoice_date)) %>%
  select(!product:state) %>%
  pivot_longer(cols = year:month,
               names_to = "date_char",
               values_to = "date_value") %>%
  pivot_longer(cols = price_per_unit:units_sold, 
               names_to = "num_char",
               values_to = "num_value") %>%
  group_by(date_char, date_value, num_char) %>%
  summarise(num_value = if_else(num_char == "price_per_unit", mean(num_value), sum(num_value))) %>%
  unique() %>%
  ungroup() %>%
  group_by(date_char) %>%
  nest()

```  


```{r}
output <- list()

for(y in 1:3){
plot <- data_time %>%
  pluck("data", y) %>%
  ggplot(mapping = aes(x = as.factor(date_value), y = num_value)) + 
    geom_col() + 
    coord_flip() +
    facet_wrap(~num_char, ncol = 3, scales = "free_x") 

output[[y]] <- plot
}

print(output)
```

Over time:  
  - More expensive things (inc mean price per unit) sold in December and July  
  - Most units sold in January, August and September  
  - More expensive items sold in 2020 but more items in total sold in 2021
  


## Bivariate descriptives

How do products sell based on multiple variables?


Set up dataset to easily check this
```{r}

data_long <- data %>%
  mutate(product_abbrev = case_when(product == "Men's Apparel" ~ "MA",
                                    product == "Men's Athletic Footwear" ~ "MAF",
                                    product == "Men's Street Footwear" ~ "MSF",
                                    product == "Women's Apparel" ~ "WA",
                                    product == "Women's Athletic Footwear" ~ "WAF",
                                    product == "Women's Street Footwear" ~ "WSF")) %>%
  mutate(year = as.character(year(invoice_date)),
         month = as.character(month(invoice_date))) %>%
  select(!c(product, state)) %>%
  pivot_longer(cols = where(is.character)) 


```

Visualize bivariate descriptives

Here, I'm looking at each categorical variable (minus state) crossed with each other.
Many plots, so I won't have them all show here.
```{r, message = FALSE}

variable <- unique(data_long$name)

for(x in variable){
  
  for(y in variable){
  
    if(x != y){
    data_plot <- data_long %>%
      filter(name == x)
    
    data_plot2 <- data_long %>%
      filter(name == y) %>%
      rename(name2 = name,
             value2 = value)
    
    data_plot_final <- full_join(data_plot, data_plot2)
    
    plot <-
      ggplot(data_plot_final, mapping = aes(x = value, y = units_sold, fill = value2)) +
      geom_boxplot() +
      theme_bw() +
      guides(fill = "none") +
      theme(axis.text.x = element_text(angle = 90)) +
      facet_wrap(~value2)
    
    print(plot)
    }
  
  }
}

```




```{r}
f <- function(x, y) {
  tbl <- data %>%
    select(x, y) %>%
    table()
  
  chisq_pval <-  round(chisq.test(tbl)$p.value, 4)
  
  data.frame(x, y, chisq_pval)
}

# create unique combinations of column names
# sorting will help getting a better plot (upper triangular)
df_comb = data.frame(t(combn(sort(names(data)), 2)), stringsAsFactors = F)

# apply function to each variable combination
df_res = map2_df(df_comb$X1, df_comb$X2, f)

# plot results
df_res %>%
  ggplot(aes(x,y,fill=chisq_pval))+
  geom_tile()+
  scale_fill_gradient(low="red", high="yellow")+
  theme_classic()


data_cor <- data %>%
  select(!c(id, state, invoice_date, units_sold, total_sales, product))

model.matrix(~0+., data=data_cor) %>% 
  cor(use="pairwise.complete.obs") %>% 
  ggcorrplot(show.diag=FALSE, type="lower", lab=TRUE, lab_size=2.5) 
```


# Prep data for models


```{r}

hist(data_mod$units_sold) # we have zero-inflated data so want to transform our data

data_mod <- data %>%
  mutate_if(is.character, as.factor) %>% # also want to convert characters to factors
  mutate(units_sold_log = log(units_sold + 0.1)) %>%
  as.data.frame()


```


# Questions

## Question 1

What is the best selling product after statistically controlling for other variables?

```{r}

# Better to use state or to use region?

summary(mod1 <- lm(units_sold_log ~ product + retailer + sales_method + state + price_per_unit,
                   data = data_mod))

summary(mod1.1 <- lm(units_sold_log ~ product + retailer + sales_method + region + price_per_unit,
                   data = data_mod))

AIC(mod1, mod1.1)


# Interaction terms?

summary(mod1.2 <- lm(units_sold_log ~ product + retailer*region + sales_method + region + price_per_unit,
                   data = data_mod))

AIC(mod1, mod1.2)


ols_plot_resid_qq(mod1)
ols_test_correlation(mod1)
ols_plot_resid_hist(mod1)
ols_plot_resid_fit(mod1)
vif(mod1)

output <- tidy(mod1)

predict <- ggeffect(mod1.1, terms = "product")


vif(mod1)

```

