---
title: "Nike Sales Data 2020-2021"
output: 
  html_document:
    toc: true
    toc_float: true
    number_section: true
    highlight: tango
    theme: readable
author: "Jamie Kerlin"
date: "2024-07-30"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```


# Overview

## Purpose

The purpose of this document is to clean and analyze the Nike Sales Dataset 2020-2021.

## Data Citation

This is a test dataset downloaded from Kaggle on 2024-07-30 https://www.kaggle.com/datasets/pavankota2020/nike-sales-dataset?resource=download

# Setup

## Load libraries

```{r}
package_names <- c("tidyverse", "here", "janitor", "vtable", "ggimage", "lubridate")

for(x in package_names){
  if(!x %in% rownames(installed.packages())) install.packages(x)
}

library(tidyverse)
library(here)
library(janitor)
library(lubridate)
library(vtable)
library(ggimage)

```


## Load data

```{r}

data <- read_csv(here("Nike_Sales_Data",
                      "Data",
                      "Data_Raw",
                      "nike_data.csv"))

```

# Data exploration and cleaning

## Clean names

Change the column names to clean names using janitor package.

```{r}

data <- data %>%
  clean_names()

```

## Column information

First, check for basics such as missing values, the number of distinct values in each column, and the data class for each of the columns.
```{r}

# 9360 rows, 9 columns

# Check for any NAs
nas <- data %>%
  filter(!complete.cases(.))

# How many distinct values are in each of the columns?
data %>%
  map_dbl(n_distinct)

# What classes of data are in the dataset?
vtable(data)
  
  
```

Next, I will look at descriptives of those column values

What are the frequencies of the values in the columns?

```{r}

sumtable(data)

data_gg <- data %>%
  pivot_longer(cols = product:sales_method,
               names_to = "characteristic",
               values_to = "characteristic_value") %>%
  group_by(characteristic, characteristic_value) %>%
  summarise(ch_num = n())


ggplot(data_gg, mapping = aes(x = characteristic_value, y = ch_num)) +
  geom_col() +
  coord_flip() +
  facet_wrap(~characteristic, scales = "free")

data_gg_num <- data %>%
  pivot_longer(cols = price_per_unit:units_sold,
               names_to = "characteristic",
               values_to = "characteristic_value")


ggplot(data_gg_num, mapping = aes(x = characteristic_value)) +
  geom_histogram() +
  facet_wrap(~characteristic, scales = "free")


data %>%
  mutate(state = fct_infreq(state)) %>%
  ggplot(mapping = aes(x = state, fill = region)) +
    geom_bar() +
    coord_flip() +
    facet_wrap(~region, scales = "free")


```

Check invoice date

I want to make sure the date format is consistent in the invoice date column

```{r}

date_test <- data %>%
  mutate(year = str_sub(invoice_date, 7, 10),
         month = str_sub(invoice_date, 4, 5),
         day = str_sub(invoice_date, 1, 2))

date_test %>% count(year)
date_test %>% count(month)
date_test %>% count(day)

# ok, date formats are correct- use lubridate on the column to get final consistent format
data$invoice_date <- lubridate::dmy(data$invoice_date)
```


Create a dataset to compare sales by categories

```{r}

data_sales <- data %>%
  pivot_longer(cols = product:state,
               names_to = "cat_char",
               values_to = "cat_value") %>%
  pivot_longer(cols = price_per_unit:units_sold, 
               names_to = "num_char",
               values_to = "num_value") %>%
  group_by(cat_char, cat_value, num_char) %>%
  summarise(mean_num_value = mean(num_value),
            total_num_value = sum(num_value)) %>%
  filter(cat_char != "state") %>%
  ungroup() %>%
  group_by(cat_char) %>%
  nest()


```


Compare average daily sales

```{r}
output <- list()

for(y in 1:4){
plot <- data_sales %>%
  pluck("data", y) %>%
  ggplot(mapping = aes(x = cat_value, y = mean_num_value)) + 
    geom_col() + 
    coord_flip() +
    facet_wrap(~num_char, ncol = 3, scales = "free_x") 

output[[y]] <- plot
}

print(output)
```

On average per day:  
  - Women and men's apparel is more expensive than footwear   
  - Women's apparel and men's street footwear sells the most units, women's athletic footwear sells the least  
  - Price per unit is lower in the south and midwest  
  - Most units sold in the west, southeast, and south  
  - Price per unit between retailers not very much  
  - Walmart sells the most units, Amazon sells the least  
  - Outlets have lower price per units  
  - Most sales from in-person  


Compare total sales over 2 years

```{r}
output <- list()

for(y in 1:4){
plot <- data_sales %>%
  pluck("data", y) %>%
  ggplot(mapping = aes(x = cat_value, y = total_num_value)) + 
    geom_col() + 
    coord_flip() +
    facet_wrap(~num_char, ncol = 3, scales = "free_x") 

output[[y]] <- plot
}

print(output)
```
Total throughout the two years:
  - Women and men's apparel is more expensive than footwear   
  - Women's apparel and men's street footwear sells the most units, followed closely by men's athletic footwear
  - Women's athletic footwear and men's apparel sell the fewest units
  - Price per unit is way higher in west and northeast
  - Most units sold in the west, northeast, and south  
  - Price per unit very different but because more units sold at some places
  - Most units sold at West Gear, Foot Locker, and Sports Direct
  - Most units sold online, fewest in-store
  
  
What do the sales look like throughout the two years?

Make a dataset to look at sales by year, quarter, month
```{r}

data_time <- data %>%
  mutate(year = year(invoice_date),
         quarter = quarter(invoice_date),
         month = month(invoice_date)) %>%
  select(!product:state) %>%
  pivot_longer(cols = year:month,
               names_to = "date_char",
               values_to = "date_value") %>%
  pivot_longer(cols = price_per_unit:units_sold, 
               names_to = "num_char",
               values_to = "num_value") %>%
  group_by(date_char, date_value, num_char) %>%
  summarise(num_value = if_else(num_char == "price_per_unit", mean(num_value), sum(num_value))) %>%
  unique() %>%
  ungroup() %>%
  group_by(date_char) %>%
  nest()

```  


```{r}
output <- list()

for(y in 1:3){
plot <- data_time %>%
  pluck("data", y) %>%
  ggplot(mapping = aes(x = as.factor(date_value), y = num_value)) + 
    geom_col() + 
    coord_flip() +
    facet_wrap(~num_char, ncol = 3, scales = "free_x") 

output[[y]] <- plot
}

print(output)
```

Over time:  
  - More expensive things (inc mean price per unit) sold in December and July  
  - Most units sold in January, August and September  
  - More expensive items sold in 2020 but more items in total sold in 2021
  
  
  